Copulas and their potential for ecology

### 2019-12-20 Paper.Rmd
refs checked

\begin{figure}[!h]
\begin{center}
\includegraphics[width=3.5in]{./Results/PedagogFig1.pdf}
\caption{(A-D) Bivariate datasets showing diverse relationships between 
variables. Some of these datasets nevertheless have the same Pearson (P)
or Spearman (S) correlation coefficients. C and D show normalized ranks of A, B, 
respectively. (E,F) Reflections
of C, D, respectively, about a vertical axis. Subscripts h and v 
in the axis labels stand for "horizontal" and "vertical".\label{fig_cop_pedag1}}
\end{center}
\end{figure}

@@

Model selection methods give the *relative* support of several models, but do not 
indicate whether any of the models are an objectively good fit. To test that, 
we tested the goodness of fit of our AIC-best copula family using a 
bootstrapping procedure of @Wang2000 and @Genest2006,
implemented as `BiCopGofTest` in `VineCopula`. The procedure performed 
one test based on a Cramer-von Mises statistic and another based on a 
Kolmogorov-Smirnov statistic. To keep computation times reasonable, 
a run using $100$ bootstraps was performed. If
the $p$-value from either test was $<0.2$, tests were re-run with 
$1000$ bootstraps.

We fit 16 bivariate copula families, exhibiting a variety of 
lower- and upper-tail dependence characteristics, with bivariate datasets. The purpose
of using a large collection of families was to include a 
variety of alternative dependence structures to have a robust model
selection and multi-model inference procedure. For that purpose, it 
is not important for the reader to
understand the details of these copulas,
and, additionally, these copulas have been 
described in detail elsewhere [@Joe1997; @Brechmann2013]. So 
we only briefly identify each family, say a few words about its tail 
dependence, and provide pictorial descriptions (Figs 
\ref{figped_N}-\ref{figped_BB1}, \ref{SM-figped_SC} and \ref{SM-figped_G}-\ref{SM-figped_SBB8}). 
The pictorial descriptions can also be used to aid quick comparisons of copula
families when alternative families are being considered for future applications. 
We used several families that can exhibit positive lower-tail dependence 
(of strength depending on parameters) and 
zero upper-tail dependence: the Clayton, survival Gumbel, survival Joe
and survival BB6 families. We used several families that can exhibit positive
upper-tail dependence (strength depends on parameters) and zero lower-tail
dependence: the survival Clayton, Gumbel, Joe and BB6 families. We used
families that show zero upper- and lower-tail dependence: the normal
and Frank families. These families have pdfs symmetric about the line
$v=-u+1$. We also used several families that can show both upper- and
lower-tail dependence, in relative amounts depending on parameter
values: the BB1, survival BB1, BB7 and survival BB7 families. 
We also used the BB8 family, which shows zero lower-tail dependence,
and zero upper-tail dependence except for a boundary case for the parameters.
And we used the survival BB8 copula, which shows zero upper-tail dependence, 
and zero lower-tail dependence except for a boundary case.
"Survival" families are rotations of the copula with the similar name by
180 degrees. 
See @Joe1997 and @Brechmann2013 for details on all families. We used the implementations
of provided in the `VineCopula` package for the R programming
language. See Figs \ref{figped_N}-\ref{figped_BB1}, \ref{SM-figped_SC} and \ref{SM-figped_G}-\ref{SM-figped_SBB8} 
for visual depictions of the pdfs of these copulas and how
pdfs and tail dependence are influenced by the parameters. See 
Table \ref{tab_copula_info} for a summary.

@@

<!--Plots for AR(1) with Clayton-->
\begin{figure}[!h]
\begin{center}
\includegraphics[width=13 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Clayton/common_legend_cause4copula_stat.pdf}\\
\vspace{-0.4 cm}
\includegraphics[width=4.3 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Clayton/Clayton_Spearman_vs_Kendall's Tau.pdf}
\includegraphics[width=4.3 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Clayton/Clayton_Kendall_vs_Kendall's Tau.pdf}
\includegraphics[width=4.3 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Clayton/Clayton_Pearson_vs_Kendall's Tau.pdf}\\
\vspace{-0.2 cm}
\includegraphics[width=4.3 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Clayton/Clayton_Corl_vs_Kendall's Tau.pdf}
\includegraphics[width=4.3 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Clayton/Clayton_Coru_vs_Kendall's Tau.pdf}
\includegraphics[width=4.3 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Clayton/Clayton_Corl-Coru_vs_Kendall's Tau.pdf}\\
\vspace{-0.2 cm}
\includegraphics[width=4.3 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Clayton/Clayton_Pl_vs_Kendall's Tau.pdf}
\includegraphics[width=4.3 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Clayton/Clayton_Pu_vs_Kendall's Tau.pdf}
\includegraphics[width=4.3 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Clayton/Clayton_Pl-Pu_vs_Kendall's Tau.pdf}\\
\vspace{-0.2 cm}
\includegraphics[width=4.3 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Clayton/Clayton_D2u_vs_Kendall's Tau.pdf}
\includegraphics[width=4.3 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Clayton/Clayton_D2l_vs_Kendall's Tau.pdf}
\includegraphics[width=4.3 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Clayton/Clayton_D2u-D2l_vs_Kendall's Tau.pdf} \\
\hspace{-0.6 cm}
\includegraphics[width=4.3 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Clayton/Clayton_scatter_CorlmCoru_Kendall's Tau.pdf}
\includegraphics[width=4.3 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Clayton/Clayton_scatter_PlmPu_Kendall's Tau.pdf}
\includegraphics[width=4.3 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Clayton/Clayton_scatter_D2umD2l_Kendall's Tau.pdf}
\caption{Comparison of correlation and tail association statistics between 
environmental-noise inputs and ecological-variable 
outputs, AR(1) model, Clayton copula. Grey and black points give 
means over 50 replicate simulations of the listed statistics, 
grey points are for noise inputs and black points are for
model outputs. Error bars are standard errors. 
$P$-values (triangles, right axis) are for a 
paired $t$-test of the null hypothesis that the distributions have the same mean. 
For bottom panels, headers give Pearson correlations for
the points. The regression line through the points (solid) was similar
to the 1-1 line (dashed). 
See section \ref{Causes} for details.\label{fig_Cause4copula_C_kend_ar1}}
\end{center}
\end{figure}

@@

\noindent Empirical results were consistent with the hypothesis that the skewness
of a spatial-average time series is influenced by tail associations
between the local quantities being averaged. 
For datasets that
exhibited highly asymmetric tail associations in earlier analyses (green spruce 
aphid abundance data had stronger lower-tail association, and, 
respectively,  leaf-curling plum 
aphid first flight data had stronger upper-tail association, Table 
\ref{tab_multivar_summary}), skewness of the spatial average 
was less than (respectively, greater than) a significant fraction of 
surrogate skewness values 
(Fig. \ref{fig_skewness_spearman}A,B). For datasets with moderately stronger 
lower than upper-tail 
dependence (\emph{Ceratium furca} abundance and methane data), 
skewness of the spatial
average showed a non-significant or marginally significant tendency 
toward being less than surrogate skewnesses (Fig. \ref{fig_skewness_spearman}C,D). 
Thus copula 
structure and asymmetric tail associations are important for spatially averaged 
quantities and can influence the temporal stability of those quantities. 
This result is represented in Fig. \ref{MasterFigure}
as the solid box around "Instability/skewness of mean or total 
time series" and the solid arrows labeled "X".

@@

Also consistent with hypothesis, copula structure had a substantial effect on
Taylor's law for the models we considered. 
Taylor's law was strongly influenced, and was often even invalidated in its assumptions of 
linearity and homoskedasticity, by non-normal copula structure. For normal copula structure 
(i.e., for simulations that gave normal-copula dependence between populations
in different locations), $p$-values for the linearity
and homoskedasticity tests were roughly uniformly distributed across replicate 
simulations and Taylor's law appeared visually to be a reasonable approximation
of the $\log(v_g)$ versus $\log(m_g)$ relationship (Fig. \ref{fig_spTL}B,E,F). 
Furthermore, quadratic coefficients
and curvature values were close to $0$ (Fig. \ref{fig_spTL}J,K), and root mean squared
errors from the linear regression were relatively small (Fig. \ref{fig_spTL}G).
But linearity or homoskedasticity were violated more frequently for non-normal 
copula structure
(Fig. \ref{fig_spTL}A,C,D,E,F); quadratic coefficients and curvatures were frequently 
non-zero (Fig. \ref{fig_spTL}J,K); and root mean squared
errors from the linear regression were much higher (Fig. \ref{fig_spTL}G). Slopes 
and intercepts of the linear regression were also strongly affected by copula
structure (Fig. \ref{fig_spTL}H,I), though some of the effect here was because
linear regressions do not always adequately represent the $\log(v_g)$ versus $\log(m_g)$
relationship when copula structure was not normal.
Thus our results substantiated the hypothesis, at least for the models we used,
that Taylor's law can be influenced by copula structure and asymmetric
tail associations. This is represented in Fig. \ref{MasterFigure} by the 
arrow labelled "Z".

@@

<!--a para on how likely one is to get copula structure via the Liebig mechanism-->
Our first proposed causal mechanism (Fig. \ref{MasterFigure}, A) may well operate commonly,
for two reasons.
First, Liebig's law and the idea of limiting 
nutrients are dominant paradigms in ecology, and 
many studies have documented nonlinear or 
threshold influences of environmental
variables on ecological quantities. 
Second, fluctuations in environmental variables through time
are very commonly correlated across space. 
Because these factors, which are the essential ingredients of
the mechanism, are common, it is reasonable to hypothesize that the mechanism
may operate commonly and may be a dominant cause of 
asymmetric tail associations and non-normal copula struture of ecological 
dependencies across space.
We provide further support for the mechanism 
in our discussion of green spruce aphids and
winter temperature below.

@@

<!--Holder stuff-->
Our simulations of character evolution suggest the hypothesis that
changes through evolutionary time in bird and 
mammal BMR and body size may exhibit greater right- than left-tail association,
contrary to standard normality assumptions of character evolution models. This is
a hypothesis only, because the greater right-tail association shown in 
Fig. \ref{fig_biv_multi_raw_cop}F,G
could have come about in another, unknown way rather than via the mechanism we suggested
which implicated asymmetric tail associations in evolutionary change. 
Our simulations show that asymmetric tail associations 
in evolutionary changes are sufficient, but may not be necessary, to produce 
the observed asymmetric tail
associations in characters of extant species. For instance (see below, and Appendix
\ref{SM-sect:missingdata} and Fig. \ref{SM-fig:missingdata}), 
systematically missing data can also produce tail dependence and may
have influenced results for the BMR-body mass datasets. Even if the hypothesized 
evolutionary mechanism (Fig. \ref{MasterFigure}, D)
is correct, our results only replace one question, i.e., why do we see greater right-
than left-tail
association between BMR and body mass, with another,
i.e., why might we see greater right- than left-tail association in evolutionary 
changes in these traits?

Performing statistical tests for associations between continuous traits across different species
    was a primary motivating example of phylogenetic comparative methods.
For example, Felsenstein's method of phylogenetically independent contrasts 
    [@Felsenstein1985] is currently the second most-cited paper in the history
    of The American Naturalist journal [@HueyGT2019].
Frustratingly, the field still lacks a dependable procedure for dealing with branch lengths, 
    which are a crucial input to the method.
Ideally, the branch lengths used to correct for phylogenetic effects would represent
    the expected amounts of change for the characters that are being analyzed.
Because researchers almost never have a reliable method for providing such branch lengths,
    most researchers rely on ultrametric trees -- those for which 
    the duration of the branch in time can be treated as a proxy for the branch length.
Frequently these branch lengths are transformed to assess sensitivity to different 
    assumptions about the degree of phylogenetic inertia displayed by the traits
    under study [@Ives2018;@Harmon2018].
Even if one were able to simply use a time-based set of branch lengths, assigning
    dates to nodes in phylogenies is difficult.
DNA sequence data can provide estimates of branch lengths, but these estimates
    are dependent on the adequacy of models which correct sequences for multiple
    substitutions occurring at the same location.
Biases in estimating the evolutionary distance can affect downstream analyses [@Phillips2009].
Additionally, changes
    in the rate of molecular evolution make the estimation of dates difficult [@HeathM2014] even when branch lengths in time are accurately estimated.

@@

<!--green spruce aphid all the way from causes to consequences-->
Our hypotheses and results cover the presence, causes, and consequences
of non-normal copula structure and asymmetric tail associations in ecological 
systems (Fig. \ref{MasterFigure}), but this was done using a variety of 
datasets and models.
We here take a closer look at the green spruce aphid, because it simultaneously 
illustrates causes and consequences within one system.
Green spruce aphid abundance, as measured in the data we use,
is strongly positively associated with the temperature of the previous winter (@Sheppard2016,
their supplementary figure 6).
Winter temperature for year $t$ was here taken to be an average for December of year $t-1$ through
March of year $t$, was available for the locations of aphid sampling, 
and was preprocessed in the same way as @Sheppard2016.
For each of our 10 sampling locations, we therefore 
examined the copula of winter temperature
and aphid abundance time series for the location, finding 
stronger left- than right-tail associations in 7 of the 8 locations for which 
independence of winter temperature and aphid abundance could be rejected, according to 
the data specific to the location (Table \ref{tab:greenspruce}).
Apparently winter temperature has an asymmetric
influence on aphid abundance in that cold winters generally produce low abundances
but warm winters often do not yield higher abundances than moderate winters.
One of our hypothesized mechanisms (Fig. \ref{MasterFigure}, A), 
which our modelling results
supported (section \ref{Results_Q2}), therefore suggests that spatial 
dependence between green spruce aphid counts
in different locations should show stronger left- than right-tail associations. This is 
exactly what was observed (Table \ref{tab_multivar_summary}), providing 
empirical evidence supporting the mechanism (this is why the box around "Nonlinear 
environmental effects, Liebig's law" and some of the arrows labelled "A" are solid
instead of dashed in Fig. \ref{MasterFigure}). The consequences of such tail 
dependence for the skewness of spatially averaged aphid counts was 
described previously 
(Fig. \ref{fig_skewness_spearman}A). Thus the asymmetric influence of winter 
temperature ultimately causes spatially averaged aphid abundance time series 
to have lower skewness (i.e., less spikiness, and greater stability through time) 
than they would otherwise. It seems likely that asymmetric influence of winter 
termperature on populations may be a common phenomenon, so effects such as 
we have documented for green spruce aphid may be common.

@@

<!--Partial Person correlation-->
We carried out most of our analyses on data ranks, $u_i$ and $v_i$, for good reasons 
mentioned in sections \ref{Intro}, \ref{Background} and \ref{Methods} and 
reviewed here; but under some circumstances it may also be appropriate to use 
techniques which parallel our nonparametric statistics but that use unranked data, 
$x_i$, $y_i$. As mentioned previously, @Genest2007 and others
recommend carrying out inferences about dependence structures (which was our
goal) using normalized ranks, stating explicitly that "statistical inference concerning
dependence structures should always be based on ranks." The reason for this is
that measures of dependence which use unranked data conflate information about
the marginal distributions of variables with information about the association
between variables. For instance, suppose the true population densities $p_{1,t}$
and $p_{2,t}$ of two species of fish in a lake in year $t$ are unknown,
but are assessed using catch per unit effort (CPUE), a standard approach in fisheries
science [@Zale2013]. The CPUE measurements, which we denote $c_{1,t}$ and $c_{2,t}$, may be 
differently correlated through time than the true densities would be, if they were known,
if the function $f$ relating $p_{i,t}$ and $c_{i,t}$ is nonlinear and if Pearson
correlation is used. The function $f$ modifies the marginal distributions, which modifies
measures such as Pearson correlation that conflate dependence and marginal information.
Rank-based measures such as Spearman or Kendall correlation, or any of our copula approaches,
will be unaffected by nonlinear monotonic functions such as $f$. The same difficulty will pertain in any
case for which a measurement of an ecological quantity is a nonlinear index of the true quantity
of interest. Nevertheless, if one is certain that measurements are linearly related to 
the true quantities being measured, statistics based on unranked data may be more appropriate 
under some circumstances, in part because ecological mechanisms which produce relationships 
between variables are influenced not only by ranks but also 
by the relative-size information in the unranked data. Judgements of what statistics
are appropriate depend, of course, on the purpose of the analysis. The goal of this 
study was to study dependence in isolation, so we used ranks. If one is interested
in tail associations using unranked data, a partial Pearson correlation can be developed
straightforwardly, by replacing $u$ and $v$ in equation \ref{eq:partialspearman}
by $x$ and $y$, but still computing the sum in that equation over data points
with ranks $u_i$ and $v_i$ constrained by the bounds $u_i+v_i>2l_b$ and 
$u_i+v_i<2u_b$. This would result in a different approach to
tail associations that would conflate dependence and marginal information but 
that may also have its own utility for some applications.

>>>>

\begin{figure}[!h]
\begin{center}
\includegraphics[width=3.5in]{./Results/PedagogFig1.pdf}
\caption{(A-D) Bivariate datasets showing diverse relationships between 
variables. Some of these datasets nevertheless have the same Pearson (P)
or Spearman (S) correlation coefficients. C and D show normalized ranks of A, B, 
respectively. (E, F) Reflections
of C, D, respectively, about a vertical axis. Subscripts h and v 
in the axis labels stand for "horizontal" and "vertical".\label{fig_cop_pedag1}}
\end{center}
\end{figure}

@@

Model selection methods give the *relative* support of several models, but do not 
indicate whether any of the models are an objectively good fit. To test that, 
we tested the goodness of fit of our AIC-best copula family using a 
bootstrapping procedure of @Wang2000 and @genest2006goodness,
implemented as `BiCopGofTest` in `VineCopula`. The procedure performed 
one test based on a Cramer-von Mises statistic and another based on a 
Kolmogorov-Smirnov statistic. To keep computation times reasonable, 
a run using $100$ bootstraps was performed. If
the $p$-value from either test was $<0.2$, tests were re-run with 
$1000$ bootstraps.

We fit 16 bivariate copula families, exhibiting a variety of 
lower- and upper-tail dependence characteristics, with bivariate datasets. The purpose
of using a large collection of families was to include a 
variety of alternative dependence structures to have a robust model
selection and multi-model inference procedure. For that purpose, it 
is not important for the reader to
understand the details of these copulas,
and, additionally, these copulas have been 
described in detail elsewhere [@joe1997multivariate; @Brechmann2013]. So 
we only briefly identify each family, say a few words about its tail 
dependence, and provide pictorial descriptions (Figs 
\ref{figped_N}-\ref{figped_BB1}, \ref{SM-figped_SC} and \ref{SM-figped_G}-\ref{SM-figped_SBB8}). 
The pictorial descriptions can also be used to aid quick comparisons of copula
families when alternative families are being considered for future applications. 
We used several families that can exhibit positive lower-tail dependence 
(of strength depending on parameters) and 
zero upper-tail dependence: the Clayton, survival Gumbel, survival Joe
and survival BB6 families. We used several families that can exhibit positive
upper-tail dependence (strength depends on parameters) and zero lower-tail
dependence: the survival Clayton, Gumbel, Joe and BB6 families. We used
families that show zero upper- and lower-tail dependence: the normal
and Frank families. These families have pdfs symmetric about the line
$v=-u+1$. We also used several families that can show both upper- and
lower-tail dependence, in relative amounts depending on parameter
values: the BB1, survival BB1, BB7 and survival BB7 families. 
We also used the BB8 family, which shows zero lower-tail dependence,
and zero upper-tail dependence except for a boundary case for the parameters.
And we used the survival BB8 copula, which shows zero upper-tail dependence, 
and zero lower-tail dependence except for a boundary case.
"Survival" families are rotations of the copula with the similar name by
180 degrees. 
See @joe1997multivariate and @Brechmann2013 for details on all families. We used the implementations
of provided in the `VineCopula` package for the R programming
language. See Figs \ref{figped_N}-\ref{figped_BB1}, \ref{SM-figped_SC} and \ref{SM-figped_G}-\ref{SM-figped_SBB8} 
for visual depictions of the pdfs of these copulas and how
pdfs and tail dependence are influenced by the parameters. See 
Table \ref{tab_copula_info} for a summary.

@@

<!--Plots for AR(1) with Clayton-->
\begin{figure}[!h]
\begin{center}
\includegraphics[width=13 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Clayton/common_legend_cause4copula_stat.pdf}\\
\vspace{-0.4 cm}
\includegraphics[width=4.3 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Clayton/Clayton_Spearman_vs_Kendall's Tau.pdf}
\includegraphics[width=4.3 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Clayton/Clayton_Kendall_vs_Kendall's Tau.pdf}
\includegraphics[width=4.3 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Clayton/Clayton_Pearson_vs_Kendall's Tau.pdf}\\
\vspace{-0.2 cm}
\includegraphics[width=4.3 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Clayton/Clayton_Corl_vs_Kendall's Tau.pdf}
\includegraphics[width=4.3 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Clayton/Clayton_Coru_vs_Kendall's Tau.pdf}
\includegraphics[width=4.3 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Clayton/Clayton_Corl-Coru_vs_Kendall's Tau.pdf}\\
\vspace{-0.2 cm}
\includegraphics[width=4.3 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Clayton/Clayton_Pl_vs_Kendall's Tau.pdf}
\includegraphics[width=4.3 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Clayton/Clayton_Pu_vs_Kendall's Tau.pdf}
\includegraphics[width=4.3 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Clayton/Clayton_Pl-Pu_vs_Kendall's Tau.pdf}\\
\vspace{-0.2 cm}
\includegraphics[width=4.3 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Clayton/Clayton_D2u_vs_Kendall's Tau.pdf}
\includegraphics[width=4.3 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Clayton/Clayton_D2l_vs_Kendall's Tau.pdf}
\includegraphics[width=4.3 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Clayton/Clayton_D2u-D2l_vs_Kendall's Tau.pdf} \\
\hspace{-0.6 cm}
\includegraphics[width=4.3 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Clayton/Clayton_scatter_CorlmCoru_Kendall's Tau.pdf}
\includegraphics[width=4.3 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Clayton/Clayton_scatter_PlmPu_Kendall's Tau.pdf}
\includegraphics[width=4.3 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Clayton/Clayton_scatter_D2umD2l_Kendall's Tau.pdf}
\caption{Comparison of correlation and tail association statistics between 
environmental-noise inputs and ecological-variable 
outputs, AR(1) model, Clayton copula. Grey and black points give 
means over 50 replicate simulations of the listed statistics, 
grey points are for noise inputs and black points are for
model outputs. Error bars are standard errors. 
$p$-values (triangles, right axis) are for a 
paired $t$-test of the null hypothesis that the distributions have the same mean. 
For bottom panels, headers give Pearson correlations for
the points. The regression line through the points (solid) was similar
to the 1-1 line (dashed). 
See section \ref{Causes} for details.\label{fig_Cause4copula_C_kend_ar1}}
\end{center}
\end{figure}

@@

\noindent Empirical results were consistent with the hypothesis that the skewness
of a spatial-average time series is influenced by tail associations
between the local quantities being averaged. 
For datasets that
exhibited highly asymmetric tail associations in earlier analyses (green spruce 
aphid abundance data had stronger lower-tail association, and, 
respectively,  leaf-curling plum 
aphid first flight data had stronger upper-tail association, Table 
\ref{tab_multivar_summary}), skewness of the spatial average 
was less than (respectively, greater than) a significant fraction of 
surrogate skewness values 
(Fig. \ref{fig_skewness_spearman}A, B). For datasets with moderately stronger 
lower than upper-tail 
dependence (\emph{Ceratium furca} abundance and methane data), 
skewness of the spatial
average showed a non-significant or marginally significant tendency 
toward being less than surrogate skewnesses (Fig. \ref{fig_skewness_spearman}C, D). 
Thus copula 
structure and asymmetric tail associations are important for spatially averaged 
quantities and can influence the temporal stability of those quantities. 
This result is represented in Fig. \ref{MasterFigure}
as the solid box around "Instability/skewness of mean or total 
time series" and the solid arrows labeled "X".

@@

Also consistent with hypothesis, copula structure had a substantial effect on
Taylor's law for the models we considered. 
Taylor's law was strongly influenced, and was often even invalidated in its assumptions of 
linearity and homoskedasticity, by non-normal copula structure. For normal copula structure 
(i.e., for simulations that gave normal-copula dependence between populations
in different locations), $p$-values for the linearity
and homoskedasticity tests were roughly uniformly distributed across replicate 
simulations and Taylor's law appeared visually to be a reasonable approximation
of the $\log(v_g)$ versus $\log(m_g)$ relationship (Fig. \ref{fig_spTL}B, E, F). 
Furthermore, quadratic coefficients
and curvature values were close to $0$ (Fig. \ref{fig_spTL}J, K), and root mean squared
errors from the linear regression were relatively small (Fig. \ref{fig_spTL}G).
But linearity or homoskedasticity were violated more frequently for non-normal 
copula structure
(Fig. \ref{fig_spTL}A, C, D, E, F); quadratic coefficients and curvatures were frequently 
non-zero (Fig. \ref{fig_spTL}J, K); and root mean squared
errors from the linear regression were much higher (Fig. \ref{fig_spTL}G). Slopes 
and intercepts of the linear regression were also strongly affected by copula
structure (Fig. \ref{fig_spTL}H, I), though some of the effect here was because
linear regressions do not always adequately represent the $\log(v_g)$ versus $\log(m_g)$
relationship when copula structure was not normal.
Thus our results substantiated the hypothesis, at least for the models we used,
that Taylor's law can be influenced by copula structure and asymmetric
tail associations. This is represented in Fig. \ref{MasterFigure} by the 
arrow labelled "Z".

@@

<!--a para on how likely one is to get copula structure via the Liebig mechanism-->
Our first proposed causal mechanism (Fig. \ref{MasterFigure}, A) may well operate commonly,
for two reasons.
First, Liebig's law and the idea of limiting 
nutrients are dominant paradigms in ecology, and 
many studies have documented nonlinear or 
threshold influences of environmental
variables on ecological quantities. 
Second, fluctuations in environmental variables through time
are very commonly correlated across space. 
Because these factors, which are the essential ingredients of
the mechanism, are common, it is reasonable to hypothesize that the mechanism
may operate commonly and may be a dominant cause of 
asymmetric tail associations and non-normal copula structure of ecological 
dependencies across space.
We provide further support for the mechanism 
in our discussion of green spruce aphids and
winter temperature below.

@@

<!--Holder stuff-->
Our simulations of character evolution suggest the hypothesis that
changes through evolutionary time in bird and 
mammal BMR and body size may exhibit greater right- than left-tail association,
contrary to standard normality assumptions of character evolution models. This is
a hypothesis only, because the greater right-tail association shown in 
Fig. \ref{fig_biv_multi_raw_cop}F, G
could have come about in another, unknown way rather than via the mechanism we suggested
which implicated asymmetric tail associations in evolutionary change. 
Our simulations show that asymmetric tail associations 
in evolutionary changes are sufficient, but may not be necessary, to produce 
the observed asymmetric tail
associations in characters of extant species. For instance (see below, and Appendix
\ref{SM-sect:missingdata} and Fig. \ref{SM-fig:missingdata}), 
systematically missing data can also produce tail dependence and may
have influenced results for the BMR-body mass datasets. Even if the hypothesized 
evolutionary mechanism (Fig. \ref{MasterFigure}, D)
is correct, our results only replace one question, i.e., why do we see greater right-
than left-tail
association between BMR and body mass, with another,
i.e., why might we see greater right- than left-tail association in evolutionary 
changes in these traits?

Performing statistical tests for associations between continuous traits across different species
    was a primary motivating example of phylogenetic comparative methods.
For example, Felsenstein's method of phylogenetically independent contrasts 
    [@Felsenstein1985] is currently the second most-cited paper in the history
    of The American Naturalist journal [@HueyGT2019].
Frustratingly, the field still lacks a dependable procedure for dealing with branch lengths, 
    which are a crucial input to the method.
Ideally, the branch lengths used to correct for phylogenetic effects would represent
    the expected amounts of change for the characters that are being analyzed.
Because researchers almost never have a reliable method for providing such branch lengths,
    most researchers rely on ultrametric trees -- those for which 
    the duration of the branch in time can be treated as a proxy for the branch length.
Frequently these branch lengths are transformed to assess sensitivity to different 
    assumptions about the degree of phylogenetic inertia displayed by the traits
    under study [@Ives2018; @Harmon2018].
Even if one were able to simply use a time-based set of branch lengths, assigning
    dates to nodes in phylogenies is difficult.
DNA sequence data can provide estimates of branch lengths, but these estimates
    are dependent on the adequacy of models which correct sequences for multiple
    substitutions occurring at the same location.
Biases in estimating the evolutionary distance can affect downstream analyses [@Phillips2009].
Additionally, changes
    in the rate of molecular evolution make the estimation of dates difficult [@HeathM2014] even when branch lengths in time are accurately estimated.

@@

<!--green spruce aphid all the way from causes to consequences-->
Our hypotheses and results cover the presence, causes, and consequences
of non-normal copula structure and asymmetric tail associations in ecological 
systems (Fig. \ref{MasterFigure}), but this was done using a variety of 
datasets and models.
We here take a closer look at the green spruce aphid, because it simultaneously 
illustrates causes and consequences within one system.
Green spruce aphid abundance, as measured in the data we use,
is strongly positively associated with the temperature of the previous winter (@Sheppard2016,
their supplementary figure 6).
Winter temperature for year $t$ was here taken to be an average for December of year $t-1$ through
March of year $t$, was available for the locations of aphid sampling, 
and was preprocessed in the same way as @Sheppard2016.
For each of our 10 sampling locations, we therefore 
examined the copula of winter temperature
and aphid abundance time series for the location, finding 
stronger left- than right-tail associations in 7 of the 8 locations for which 
independence of winter temperature and aphid abundance could be rejected, according to 
the data specific to the location (Table \ref{tab:greenspruce}).
Apparently winter temperature has an asymmetric
influence on aphid abundance in that cold winters generally produce low abundances
but warm winters often do not yield higher abundances than moderate winters.
One of our hypothesized mechanisms (Fig. \ref{MasterFigure}, A), 
which our modelling results
supported (section \ref{Results_Q2}), therefore suggests that spatial 
dependence between green spruce aphid counts
in different locations should show stronger left- than right-tail associations. This is 
exactly what was observed (Table \ref{tab_multivar_summary}), providing 
empirical evidence supporting the mechanism (this is why the box around "Nonlinear 
environmental effects, Liebig's law" and some of the arrows labelled "A" are solid
instead of dashed in Fig. \ref{MasterFigure}). The consequences of such tail 
dependence for the skewness of spatially averaged aphid counts was 
described previously 
(Fig. \ref{fig_skewness_spearman}A). Thus the asymmetric influence of winter 
temperature ultimately causes spatially averaged aphid abundance time series 
to have lower skewness (i.e., less spikiness, and greater stability through time) 
than they would otherwise. It seems likely that asymmetric influence of winter 
temperature on populations may be a common phenomenon, so effects such as 
we have documented for green spruce aphid may be common.

@@

<!--Partial Person correlation-->
We carried out most of our analyses on data ranks, $u_i$ and $v_i$, for good reasons 
mentioned in sections \ref{Intro}, \ref{Background} and \ref{Methods} and 
reviewed here; but under some circumstances it may also be appropriate to use 
techniques which parallel our nonparametric statistics but that use unranked data, 
$x_i$, $y_i$. As mentioned previously, @Genest2007 and others
recommend carrying out inferences about dependence structures (which was our
goal) using normalized ranks, stating explicitly that "statistical inference concerning
dependence structures should always be based on ranks". The reason for this is
that measures of dependence which use unranked data conflate information about
the marginal distributions of variables with information about the association
between variables. For instance, suppose the true population densities $p_{1,t}$
and $p_{2,t}$ of two species of fish in a lake in year $t$ are unknown,
but are assessed using catch per unit effort (CPUE), a standard approach in fisheries
science [@Zale2013]. The CPUE measurements, which we denote $c_{1,t}$ and $c_{2,t}$, may be 
differently correlated through time than the true densities would be, if they were known,
if the function $f$ relating $p_{i,t}$ and $c_{i,t}$ is nonlinear and if Pearson
correlation is used. The function $f$ modifies the marginal distributions, which modifies
measures such as Pearson correlation that conflate dependence and marginal information.
Rank-based measures such as Spearman or Kendall correlation, or any of our copula approaches,
will be unaffected by nonlinear monotonic functions such as $f$. The same difficulty will pertain in any
case for which a measurement of an ecological quantity is a nonlinear index of the true quantity
of interest. Nevertheless, if one is certain that measurements are linearly related to 
the true quantities being measured, statistics based on unranked data may be more appropriate 
under some circumstances, in part because ecological mechanisms which produce relationships 
between variables are influenced not only by ranks but also 
by the relative-size information in the unranked data. Judgements of what statistics
are appropriate depend, of course, on the purpose of the analysis. The goal of this 
study was to study dependence in isolation, so we used ranks. If one is interested
in tail associations using unranked data, a partial Pearson correlation can be developed
straightforwardly, by replacing $u$ and $v$ in equation \ref{eq:partialspearman}
by $x$ and $y$, but still computing the sum in that equation over data points
with ranks $u_i$ and $v_i$ constrained by the bounds $u_i+v_i>2l_b$ and 
$u_i+v_i<2u_b$. This would result in a different approach to
tail associations that would conflate dependence and marginal information but 
that may also have its own utility for some applications.
